---
title: "a3_task3_michelle_leporibui"
author: "Michelle Lepori-Bui"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(kableExtra)
library(lubridate)
library(here)
```
```{r}
urchins <- read_csv(here("data","urchins.csv"))
```

Section 1: size comparison, purple and red urchins
include data from all sites and for both treatments
ie ignore those variables in this exploration

i. exploratory data visualization
compare districution of sizes between purple and rd urchins using the graph types below 
these are exploratory not final so they dont need to be perfect and dont add fig captions but at least update axis labels
- faceted histogram (hint geom_histogram then facet_wrap (~common_name)
- boxplot (geom_boxplot) with common name on the x axis and size on the y axis

```{r}
urchins_clean <- urchins %>% 
  clean_names()
ggplot(data = urchins_clean, aes(x = size)) +
  geom_histogram() +
  facet_wrap(~common_name) +
  labs (x = "diameter (cm)", y = "number of urchins")

ggplot(data = urchins_clean, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs( x =  "urchin species", y = "diameter (cm)")
  
```


ii. descriptive statistics table
create a summary table that contains the following statistics for purple abd red urchins (still ignoring site as a variable): 
mean
median
std dev
sample size
kableExtra to make it nicer 
hint: group_by() %>% summarize(mean,median,sd,n)

```{r}
urchins_summary <- urchins_clean %>% 
  group_by(common_name) %>% 
  summarize(urch_mean = mean(size),
            urch_med = median(size),
            urch_se = sd(size),
            urch_n = n())
  
urchins_summary %>% 
  kable(col.names = c("species", "mean", "median", "standard deviation", "sample size"))
```

## 2 - 3 sentences here

ii. below where your exploratory graphs and table appear add 2 -3  sentences briefly describing your overall takeaways about how the size (ditributinm centrat tendency, spread, etc ) compare between the two spps of urchins 
contextualize any differences you observe - for example, what is the actuak and % differene in mean urchin diameter betewen the spps?

Sectiom 2. explore purple urchin counts by site and year
i. in a piped sequence create stored subset of the urchins data called purple_urchins in which you
filter( spps = purple)
convert date to date class
mutate new cokumn only year
new column called site_full that contaisn the full name not abbrev for each of the 5 sites:
Naples Reef
Arroyo Quemado Reef
Carpinteria REef
Mohawk REed
Isla Vista REed
hint case_when()

```{r}
purple_urchins <- urchins_clean %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ))
```


ii. summary table containing the total number of purple urchins observed by site and year 
starting from purple urchins and using count() function to find counts of obs by group
no need to finalize this table just use output in plot in next step

```{r}
purple_urchins_summary <- purple_urchins %>% 
  group_by(site, year) %>% 
  summarize(count = n()) 
purple_urchins_summary
```


iii. make an exploratory graph of counts of purple urchins recorded at each site over time for the years included inthe study not finalized

```{r}
ggplot(data = purple_urchins_summary, aes(x = year, y = count)) +
  geom_col(aes(fill = site)) +
  facet_wrap(~site)
```


iv. save and knit














